The \project{unRAVE} catalog

### 2016-08-23 article/unrave.tex
Some stubs for me to write out tomorrow

\begin{abstract}
The Milky Way may be our greatest laboratory for understanding galaxy formation.
This premise relies on the ability to measure --- for large numbers of stars ---
3D positions and motions, and their detailed chemical abundances.  On 14 September
2016 the \gaia\ mission will release the former for $\approx2\times10^6$ stars.
However, a catalog of precise chemical abundances to complement that data
release is currently lacking.  The RAdial Velocity Experiment (\rave) survey has 
acquired spectra for 292,036 stars in common with the first \gaia\ data release, 
constituting the largest overlap of any stellar spectroscopic sample.  Here we 
perform an independent analysis of \rave\ spectra using an implementation of 
\thecannon\ that uses prior probabilities on atomic line formation and stellar
astrophysical parameters.  Our approach yields precise stellar labels at lower
S/N ratios than most \emph{ab initio} approaches, allowing us to deliver
improved radial velocities, effective temperature $\teff$, surface gravity $\logg$,
and chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) 
for \Nstars\ stars.  We validate our results internally from repeat visits, 
and externally against existing high-resolution spectroscopic surveys.  The 
typical precision in chemical abundances is X.XX~dex.  We provide 
spectrophotometric distances based on our labels, improving upon the parallax
distances available for the fainter targets in the first \gaia\ data release.
\stub{A one-line science result based on RAVE/Hipparcos overlap}
\end{abstract}

@@

% Identifying obviously bad stars that have been mis-classified by K2/EPIC

% Giant stars of low metallicity.

% Refer to the training set figure.
% Two separate models, with an overlap region.
% Worry later about label offsets.
% Make sure you mention the lack of abundance information in K2/EPIC

% Initially we train on three labels.

% Inserting 2MASS photometry as fake pixels

% Magic wizardry for missing and imperfect labels.
% Censoring masks on abundance labels
% Priors on astrophysical parameters from isochrones

% Include velocities and rotational velocities at test time?

@@

@@

\subsection{Internal validation}
We first conducted leave-one-out cross-validation to validate our results.
For each training set star we removed that star from the training set and
then re-trained the model.  Using that re-trained model, we then treated
the excluded star as a test star and compared our inferred labels to the
test set labels.  We show the results of this experiment in Figure
\ref{fig:loocv-comparison}.  The bias and root mean square (RMS) error in 
$\teff$ is XX~K and XX~K, and in $\logg$ is X.XX~dex and X.XX~dex,
respectively.  The bias in abundance labels is negligible --- between
X.XX to X.XX --- and the RMS varies from X.XX ([XX/H]) to X.XX ([XX/H]).

@@

% Identify outliers:
% --> Chi-sq of the sample.
% --> Convex hull? Distance to plane?
% --> Stars with morphological classification
% --> Anything else?

@@

\acknowledgements
A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt) and Sven Buder (MPIA).
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360. Funding for KH has been provided through the Simons Foundation Society of Fellows and the Marshall Scholarship.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.

>>>>

% Stubs for discussion & analysis.
% TODO: Inserting 2MASS photometry as fake pixels
% TODO: Censoring masks on abundance labels
% TODO: Priors on astrophysical parameters from isochrones
% TODO: Include velocities and rotational velocities at test time?

\begin{abstract}
The Milky Way may be our greatest laboratory for understanding galaxy formation.
This premise relies on the ability to measure --- for large numbers of stars ---
their detailed chemical abundances, and their 3D positions and motions.  On 14 September
2016 the \gaia\ mission will release the latter for $\approx2\times10^6$ stars.
However, a catalog of precise chemical abundances to complement that data
release is currently lacking.  The RAdial Velocity Experiment (\rave) survey has 
acquired spectra for 292,036 stars in common with the first \gaia\ data release, 
constituting the largest overlap of any stellar spectroscopic sample.  Here we 
perform an independent analysis of \rave\ spectra using an implementation of 
\thecannon\ that uses prior probabilities on atomic line formation and stellar
astrophysical parameters.  Our approach yields precise stellar labels at lower
S/N ratios than most \emph{ab initio} approaches, allowing us to deliver
improved radial velocities, effective temperature $\teff$, surface gravity $\logg$,
and chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) 
for \Nstars\ stars.  We validate our results internally from repeat visits, 
and externally against existing high-resolution spectroscopic surveys.  The 
typical precision in chemical abundances is X.XX~dex.  We provide 
spectrophotometric distances based on our labels, improving upon the parallax
distances available for the fainter targets in the first \gaia\ data release.
\stub{A one-line science result based on RAVE/Hipparcos overlap}
\end{abstract}

@@

% Identifying obviously bad stars that have been mis-classified by K2/EPIC
% Requiring stars with narrow posteriors.
% Make sure you mention the lack of abundance information in K2/EPIC

% Including Giant stars of low metallicity.
% Refer to the training set figure.
% Worry later about label offsets.

@@

% Describe The Cannon basic model, vectorizer, etc.
% Fitting time errors, etc.

% Here we will first consider only a three parameter model.
% APOGEE used only quadratic model, worked for giants
% We experimented with a quadratic model only. 
% Useful for identifying mis-classifications but the structure wasn't all there..
% worked but didn't show the structure in detail.
% Split into two different models with partial overlap.
% One for giants and one for dwarfs.
% The labels for stars in the training sets are shown in Figure X.

% We trained all three models.
% Note the differences in metallicity scales.
% The K2/EPIC stuff is accurate but imprecise, whereas the APOGEE stuff is precise, usually precise (except Ti?)
% Look at spectral derivatives.
% Show the model internals are relatable.
% Map spectral derivatives from giant model to main-sequence model.
% Update the model with those new spectral derivatives.

% We ran all three models on all spectra.
% How we combined the results into one.
% Refer to test set HRD.

% Restricted to 3-label model, but the giant set includes abundance labels.
% The spectral range covered by \rave\ includes atomic transitions of X, Y, Z elements.
% The strongest lines are indicated in Figure XXX.

% Censoring masks
% Transfering coefficients, etc.

@@

\subsection{Internal validation}
We first conducted leave-one-out cross-validation to validate our results.
For each training set star we removed that star from the training set and
then re-trained the model.  Using that re-trained model, we then treated
the excluded star as a test star and compared our inferred labels to the
test set labels.  We show the results of this experiment in Figure
\ref{fig:loocv-comparison}.  The bias and root mean square (RMS) error in 
$\teff$ is XX~K and XX~K, and in $\logg$ is X.XX~dex and X.XX~dex,
respectively.  The bias in abundance labels is negligible --- between
X.XX to X.XX --- and the RMS varies from X.XX ([XX/H]) to X.XX ([XX/H]).
% Fill this in tomorrow.

@@

% Identify outliers:
% --> Chi-sq of the sample.
% --> Convex hull? Distance to plane?
% --> Stars with morphological classification
% --> Some rogue gallery of a few high S/N cases that have clearly gone haywire?

@@

\acknowledgements
A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt), Sergey Koposov (Cambridge),
and Sven Buder (MPIA).
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360. Funding for KH has been provided through the Simons Foundation Society of Fellows and the Marshall Scholarship.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.
